---
title: "pm566-midterm-fall2022"
author: "Misha Khan"
date: "2022-10-21"
output: html_document
---

```{r}
library(stringr)
library(dplyr)
library(tidyverse)
library(rvest)
library(xml2)
library(dbplyr)
library(dplyr)
library(skimr) 
library(ggplot2)
library(data.table)
library(lubridate)
```

Topic: Crime Data from 2020 to Present in Los Angeles
Missing data (0, 0) for long and lat are for privacy
URL: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8
```{r}
crimeraw <- read.csv("Crime_Data_from_2020_to_Present.csv")
```

Data Cleaning
Rename cols to make it more meaningful
```{r}
colnames(crimeraw) <- c('Record No',
                        'Date Reported',
                        'Date Occurred',
                        'Time Occurred',
                        'Area No',
                        'Area',
                        'Rpt Dist No', #drop
                        'Part', #drop
                        'Crime Code',
                        'Crime',
                        'MO', #drop
                        'Victim Age',
                        'Victim Sex',
                        'Victim Ethnicity',
                        'Premise Code',
                        'Premise',
                        'Weapon Code',
                        'Weapon',
                        'Case Code',
                        'Case Status',
                        'Crime Code 1',
                        'Crime Code 2',
                        'Crime Code 3',
                        'Crime Code 4',
                        'Location',
                        'Cross Street',
                        'Lat',
                        'Lon'
                        )

crimeraw <- select(crimeraw, -c('Rpt Dist No','Part', 'MO','Crime Code 1', 'Crime Code 2', 'Crime Code 3', 'Crime Code 4', 'Cross Street'))
```

Reformat date variables (EDA: Date when most crimes occur)
```{r}
crimeraw$`Date Reported` <- as.Date(crimeraw$`Date Reported`, format = "%m/%d/%Y")
crimeraw$`Date Occurred` <- as.Date(crimeraw$`Date Occurred`, format = "%m/%d/%Y")
```

Change miltary to standard time (Possible EDA: Time when most crime occur)
```{r}
crimeraw$`Time Occurred` <- format(strptime(substr(as.POSIXct(sprintf("%04.0f", crimeraw$`Time Occurred`), 
                                 format="%H%M"), 12, 16), '%H:%M'), '%I:%M %p')
```

View missing data
```{r}
skimr::skim(crimeraw)
```

MISSING: Weapon Code
EMPTY: Victim Sex, Victim Ethnicity, Premise, Weapon
IF missing, change to 999
FIX: empty cells will not change to value or NA
```{r}
crimeraw$`Weapon Code` <- crimeraw$`Weapon Code` %>% replace_na(999)
```


New dataset created
START HERE
```{r}
#Save as changes as a new file
write_rds(crimeraw, "crimedatnew.rds")
crimedat <- read_rds("crimedatnew.rds")
crimedat <- as.data.frame(crimedat)
```



EDA
Extract top 10 areas
```{r}
#Create new df to arrange 
crime_frequency <-count(crimedat, Area)

#Create table of crime frequency highest to lowest
crime_frequency %>%
  arrange(desc(crime_frequency$n)) %>%
knitr::kable() 

#Filter df to only focus on top 10 areas
#crimedat <- subset(crimedat, Area %in% c("Central", "77th Street", "Pacific", "Southwest", "Hollywood", "Southeast", "N Hollywood", "Newton", "Olympic", "Wilshire"))
```

You don't want to subset the data before you've computed percentages.

EDA #1: What is the amount of crime per area in LA?
Count of crime per area

Proportions of crime in area
FIX: figure out how to combine percents into the same table
```{r}
#Percentage of Crime
sort(round(prop.table(table(crimedat$`Area`)) * 100, 3), T) %>%
knitr::kable()
```

```{r}
nc <- sum(crime_frequency$n)

crime_frequency %>% 
  arrange(desc(n)) %>%
  mutate(pct = round( n / nc * 100,1))
```


Visualization of top 10 areas
```{r}
v1 <- table(crimedat$`Area`)
barplot(sort(v1, T)[1:10], las = 2, col = rainbow(12), cex.names= .7, main = "Top 10 Areas with Crime", xlab = "Area", ylab = "Count")     
```

Get month before counting crimes by month and area.

```{r}
crimedat$Month = format(crimedat$`Date Occurred`, "%B")

# convert dates to character data
#strDates <- as.character(dates)
```

```{r}
#Create CrimeCount col for the Area
crimedat <- crimedat %>%
  group_by(Month,`Area`) %>% 
  summarize(CrimeCount = n())
```



FIX: find month with the highest crime (3 years of data)
```{r}
crimedat %>%
ggplot(aes(x = Month, y = CrimeCount, fill = Area)) +
  geom_bar(stat = "identity")
```


